---
title: "Big Mac Index"
author: "Albert Campillo"
date: "1/23/2021"
output: html_document
editor_options:
    chunk_output_type: console
---

```{r setup, include=FALSE} 

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

library(scales)
library(ggrepel)
theme_set(theme_light())
```

## Load the weekly data

Download the weekly data and make it available in the “tt“ object

```{r Load}

tt <- tt_load("2020-12-22") 
tt$`big-mac` %>% View()

big_mac <- tt$`big-mac` %>%
    rename(country = name) %>%
    
    #count nbr of entries per country (used later to exclude countries w/ missing data)
    add_count(country, name = "country_total") %>%
    
    # only countries w/ complete data
    filter(country_total == max(country_total)) %>%
    
    arrange(country, date)

```


```{r}
big_mac %>%
    
    # order countries by their inflation (function of local currency max /  min)
    mutate(country = fct_reorder(country, local_price, function(.) last(.) / first(.))) %>% 
    
    ggplot(aes(date, local_price, color = country)) +
    geom_line() +
    expand_limits(y = 0) +
    facet_wrap(~country, scales = "free_y") +
    theme(legend.position = "none") +
    labs(x = "Time",
         y = "Prices of Big Mac in local currency") 
```


```{r}
big_mac %>%
    group_by(country) %>%
    summarize(big_mac_inflation = last(local_price) / first(local_price)) %>%
    arrange(desc(big_mac_inflation)) %>%
    mutate(country = fct_reorder(country, big_mac_inflation)) %>%
    
    ggplot(aes(big_mac_inflation, country)) +
    geom_col() +
    geom_text(aes(label = paste0(round(big_mac_inflation, 1), "x")), hjust = 0) +
    scale_x_log10(breaks = c(1,3,10,30,100)) +
    expand_limits(x = 130) +
    labs(x= "Price of Big Mac in 2020 / Price of Big Mac in 2000")
    

```


```{r} 
big_mac %>%
    group_by(date) %>%
    mutate(usd_price = local_price[iso_a3== "USA"],
           usd_gdp   = gdp_dollar[iso_a3 == "USA"]) %>%
    mutate(big_mac_ex = local_price / usd_price) %>%
    select(date, iso_a3, country, local_price, dollar_ex, usd_price, big_mac_ex, usd_raw, usd_gdp, gdp_dollar, usd_adjusted) %>%
    mutate(usd_recalculated = (big_mac_ex - dollar_ex) / dollar_ex) %>%
    filter(!is.na(gdp_dollar))

```


```{r}
big_mac %>%
    select(date, country, local_price, dollar_ex, usd_raw, gdp_dollar, usd_adjusted) %>%
    filter(!is.na(gdp_dollar)) %>%
    
    ggplot(aes(date, usd_raw)) +
    geom_line() +
    expand_limits(y = 0) +
    facet_wrap(~country, scales = "free_y")
    
```


```{r}
# Why Argentinian currency is undervalued? (hint: inflation)

big_mac %>%
    group_by(date) %>%
    mutate(usd_price = local_price[iso_a3== "USA"],
           usd_gdp   = gdp_dollar[iso_a3 == "USA"]) %>%
    ungroup() %>%
    filter(country == "Argentina",
           !is.na(gdp_dollar)) %>%
    mutate(price_from_usd = usd_price * dollar_ex) %>%
     
    ggplot(aes(date, local_price)) +
    geom_line(aes(color = "Price (in local currency)")) +
    geom_line(aes(y = price_from_usd, color = "Price from USD"))


    
```


```{r}
big_mac %>%
    select(date, country, local_price, dollar_ex, usd_raw, gdp_dollar, usd_adjusted) %>%
    filter(!is.na(gdp_dollar)) %>%
    mutate(country = fct_reorder(country, usd_raw)) %>%
    
    ggplot(aes(date, usd_raw)) +
    geom_line() +
    geom_hline(color = "red", lty = 2, yintercept = 0) +
    expand_limits(y = 0) +
    facet_wrap(~country) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(y = "Big Max Index relative to USD",
         x = "")


big_mac %>%
    filter(country != "United States") %>%
    select(date, country, local_price, dollar_ex, usd_raw, gdp_dollar, usd_adjusted) %>%
    filter(!is.na(gdp_dollar)) %>%
    mutate(country = fct_reorder(country, usd_raw)) %>%
    
    ggplot(aes(date, usd_adjusted)) +
    geom_line() +
    geom_hline(color = "red", lty = 2, yintercept = 0) +
    expand_limits(y = 0) +
    facet_wrap(~country) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(y = "Raw Big Max Index relative to USD",
         x = "")

 
big_mac %>%
    filter(date == max(date)) %>%
    ggplot(aes(gdp_dollar, usd_raw)) +
    geom_point() +
    geom_smooth(method = "lm") + 
    geom_text_repel(aes(label = country)) + 
    labs(x = "GDP per capita (dollars)",
         y = "Raw Big Mac Index relative to USD")
```

In Indonesia, Big Macs are cheaper than expected based on the currency conversion rate. 
However, this is not surprising as for counties with Indonesia´s GDP per capita it's roughly par for the course. 

What we are really interested are rich countries with expensive Big Macs, which indicate that their currency is either undervalued or overvalued respectively.

```{r}
big_mac %>%
    filter(date == max(date)) %>%
    ggplot(aes(gdp_dollar, usd_adjusted)) +
    geom_point() +
    geom_smooth(method = "lm") + 
    geom_text_repel(aes(label = country)) + 
    labs(x = "GDP per capita (dollars)",
         y = "Adjusted Big Mac Index relative to USD")

```


```{r}

big_mac %>%
    filter(date == max(date)) %>%
    mutate(country = fct_reorder(country, usd_adjusted)) %>%
    
    ggplot(aes(usd_adjusted, country)) +
    geom_col() +
    labs(x = "Big Mac Index (GDP adjusted)",
         y = "")
```  

```{r}
library(gganimate)

big_mac %>%
    filter(!is.na(gdp_dollar)) %>%
    ggplot(aes(gdp_dollar, usd_adjusted)) +
    geom_point() +
    geom_smooth(method = "lm") + 
    geom_text_repel(aes(label = country)) + 
    transition_time (date) + 
    labs(x = "GDP per capita (dollars)",
         y = "Adjusted Big Mac Index relative to USD", 
         title = "{ frame_time }")


``` 

```{r}
big_mac %>%
    select(date, country, local_price, dollar_ex, usd_raw, gdp_dollar, ends_with("adjusted")) %>%
    pivot_longer(ends_with("adjusted"), names_to = "base_currency", values_to = "adjusted") %>%
    
    mutate(base_currency = str_to_upper(str_remove(base_currency, "_adjusted"))) %>%
    
    filter(!is.na(gdp_dollar)) %>%
    mutate(country = fct_reorder(country, adjusted )) %>%
    
    ggplot(aes(date, adjusted, color = base_currency)) +
    geom_line() +
    geom_hline(color = "red", lty = 2, yintercept = 0) +
    expand_limits(y = 0) +
    facet_wrap(~country) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(y = "GCP-adjusted Big Max Index relative to USD",
         x = "") 

```
