---
title: "Big Mac Index"
author: "Albert Campillo"
date: "1/23/2021"
output: html_document
editor_options:
    chunk_output_type: console
---

```{r setup, include=FALSE} 

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

library(scales)
library(ggrepel)

theme_set(theme_light())
```

## Load the weekly data

Download the weekly data and make it available in the “tt“ object

```{r Load}

tt <- tt_load("2020-12-22") 
tt$`big-mac` %>% View()

big_mac <- tt$`big-mac` %>%
    rename(country = name) %>%
    
    #count nbr of entries per country (used later to exclude countries w/ missing data)
    add_count(country, name = "country_total") %>%
    
    # only countries w/ complete data
    filter(country_total == max(country_total)) %>%
    
    arrange(country, date)

```


usd_raw -> indicate whether a currency is undervalued or overvalued vs the usd as a result of comparing the price of the local burger at usd price (local_price / usd_price) vs the usd exchange rate (usd_price), therefore:

((local_price / usd_price) - dollar_ex) / dollar_ex = undervalued (+) or overvalued(-)


```{r}
# Evolution of inflation rate by country over time

big_mac %>%
    # sort countries by their inflation evolution over time (function of local currency last / first)
    mutate(country = fct_reorder(country, local_price, function(.) last(.) / first(.))) %>% 
    
    ggplot(aes(date, local_price, color = country)) +
    geom_line() +
    expand_limits(y = 0) +
    facet_wrap(~country, scales = "free_y") +
    theme(legend.position = "none") +
    labs(x = "Time",
         y = "Prices of Big Mac in local currency") 
```


```{r}
# What inflation has each country experienced over the period?

big_mac %>%
    group_by(country) %>%
    summarize(big_mac_inflation = last(local_price) / first(local_price)) %>%
    arrange(desc(big_mac_inflation)) %>%
    mutate(country = fct_reorder(country, big_mac_inflation)) %>%
    
    ggplot(aes(big_mac_inflation, country)) +
    geom_col() +
    geom_text(aes(label = paste0("  ", round(big_mac_inflation, 1), "x")), hjust = 0) +
    scale_x_log10(breaks = c(1,3,10,30,100)) +
    expand_limits(x = 130) +
    labs(x= "Price of Big Mac in 2020 / Price of Big Mac in 2000",
         title = "Inflation multiple by country")
    

```


```{r} 
# check the input values of local currency overvalue and undervalue

big_mac %>%
    group_by(date) %>%
    mutate(usd_price = local_price[iso_a3 == "USA"],
           usd_gdp   = gdp_dollar [iso_a3 == "USA"]) %>%
    ungroup() %>%
    mutate(big_mac_ex = local_price / usd_price) %>%
    select(date, iso_a3, country, local_price, dollar_ex, usd_price, big_mac_ex, usd_raw, usd_gdp, gdp_dollar, usd_adjusted) %>%
    mutate(usd_recalculated = (big_mac_ex - dollar_ex) / dollar_ex) %>%
    filter(!is.na(gdp_dollar))

```


```{r}
# What is the adjusted Big Mac index relative to the dollar over time?

big_mac %>%
    select(date, country, local_price, dollar_ex, usd_raw, gdp_dollar, usd_adjusted) %>%
    filter(!is.na(gdp_dollar)) %>%
    
    ggplot(aes(date, usd_raw)) +
    geom_line() +
    expand_limits(y = 0) +
    facet_wrap(~country, scales = "free_y") +
    labs(x = "Period ",
         y = "Currency value relative to USD",
         z = "Big Mac Index Evolution per country")
    
```



```{r}
# Is the argentinian peso undervalued? (hint: inflation)

big_mac %>%
    group_by(date) %>%
    mutate(usd_price = local_price[iso_a3== "USA"],
           usd_gdp   = gdp_dollar[iso_a3 == "USA"]) %>%
    ungroup() %>%
    filter(country == "Argentina",
           !is.na(gdp_dollar)) %>%
    
    # calculate what would the LC price in USD be
    mutate(price_from_usd = usd_price * dollar_ex) %>%
     
    ggplot(aes(date, local_price)) +
    geom_line(aes(color = "Price (in local currency)")) +
    geom_line(aes(y = price_from_usd, color = "Price from USD"))

```

The prior chart shows that, if we compare the price of the Big Mac in pesos and the value of that big mac converted to usd, the value of the peso is much lower than the usd (we would have expected it much higher than actual=.
This implies there are other factors affecting the undervalued situation of the peso.

Hint: (Hyper)Inflation


```{r}
big_mac %>%
    select(date, country, local_price, dollar_ex, usd_raw, gdp_dollar, usd_adjusted) %>%
    filter(!is.na(gdp_dollar),
           country != "United States") %>%
    
    #order countries by most undervalued currency ones to more overvalued
    mutate(country = fct_reorder(country, usd_raw)) %>%
    
    ggplot(aes(date, usd_raw)) +
    geom_line() +
    geom_hline(color = "red", lty = 2, yintercept = 0) +
    expand_limits(y= 0 ) +
    facet_wrap(~country) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(y = "Big Mac Index relative to USD",
         x ="")


# Analyse it from the gdp perspective
# Scatterplot each country usd_raw vs its gpd_dollar

big_mac %>%
    filter(date == max(date)) %>%
    ggplot(aes(gdp_dollar, usd_raw)) +
    geom_point() +
    geom_smooth(method = "lm") + 
    geom_text(aes(label = country)) +
    labs(x = "GDP per capita (USD)",
         y = "Raw Big Mac Index relative to USD")

```

In Indonesia, Big Macs are cheaper than expected based on the currency conversion rate. 
However, this is not surprising as for countries with Indonesia´s GDP per capita it's roughly par for the course. 

What we are really interested are rich countries with cheap Big Macs,or relatively poor countries with expensive Big Macs, which indicate that their currency is either undervalued or overvalued respectively.

```{r}
# Big Mac Index in USD in 2020 accross countries 

big_mac %>%
    filter(date == max(date)) %>%
    ggplot(aes(gdp_dollar, usd_adjusted)) +
    geom_point() +
    geom_smooth(method = "lm") + 
    geom_text_repel(aes(label = country)) + 
    labs(x = "GDP per capita (dollars)",
         y = "Adjusted Big Mac Index relative to USD")

```


```{r}
# Animate previous plot

library(gganimate)

big_mac %>%
    filter(!is.na(gdp_dollar)) %>%
    ggplot(aes(gdp_dollar, usd_adjusted)) +
    geom_point() +
    geom_smooth(method = "lm") + 
    geom_text_repel(aes(label = country)) + 
    transition_time(date) + 
    labs(x = "GDP per capita (dollars)",
         y = "Adjusted Big Mac Index relative to USD", 
         title = "{ frame_time }")
```


```{r}

# Which countries have most undervalued currencies?

big_mac %>%
    filter(date == max(date)) %>%
    mutate(country = fct_reorder(country, usd_adjusted)) %>%
    
    ggplot(aes(usd_adjusted, country)) +
    geom_col() +
    labs(x = "Big Mac Index (GDP adjusted)",
         y = "",
         title = "Rank of countries by currency value (2020)",
         subtitle = "From most undervalued to most overvalued")
```  


```{r}
big_mac %>%
    select(date, country, local_price, dollar_ex, usd_raw, gdp_dollar, 
           ends_with("adjusted")) %>%
    
    pivot_longer(ends_with("adjusted"), names_to = "base_currency", values_to = "adjusted") %>%
    
    mutate(base_currency = str_to_upper(str_remove(base_currency, "_adjusted"))) %>%
    
    filter(!is.na(gdp_dollar)) %>%
    mutate(country = fct_reorder(country, adjusted)) %>%
    
    ggplot(aes(date, adjusted, color = base_currency)) +
    geom_line() +
    geom_hline(color = "red", lty = 2, yintercept = 0) +
    expand_limits(y = 0) +
    facet_wrap(~country) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(y = "GDP-Adjusted Big Max Index relative to each currency",
         x = "") 

```



What Big Mac prices tend to move together with the USD?

```{r}
library(widyr)

big_mac %>%
    pairwise_cor(country, date, local_price, sort = TRUE) %>%
    filter(item1 == "United States") %>%
    View()
``` 